---
title: "Brand & Model Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: cosmo
---

```{r setup, include=FALSE}
library(tidyverse)
library(scales)
library(flexdashboard)
library(knitr)
library(tidytext)
```

```{r}
# Les inn data og gjør klar
df <- read_csv("brand_model_data.csv") %>%
  filter(!is.na(model), model != "", toupper(model) != "OTHERS") %>%
  mutate(
    country_full = if_else(country_code == 0, "Denmark", country_full),
    country_full = str_to_title(country_full)
  ) %>%
  group_by(country_full, brand, model) %>%
  summarise(num_ads = sum(num_ads), .groups = "drop")

# Topp 8 land og 10 merker
top_countries <- df %>%
  group_by(country_full) %>%
  summarise(total = sum(num_ads), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 8) %>%
  pull(country_full)

top_brands <- df %>%
  group_by(brand) %>%
  summarise(total = sum(num_ads), .groups = "drop") %>%
  arrange(desc(total)) %>%
  slice_head(n = 10) %>%
  pull(brand)
```

Column {data-width=700}
-----------------------------------------------------------------------

### Top 10 Brands – Market Share by Country

```{r}
df_brand_country <- df %>%
  filter(country_full %in% top_countries, brand %in% top_brands) %>%
  group_by(country_full, brand) %>%
  summarise(num_ads = sum(num_ads), .groups = "drop") %>%
  group_by(country_full) %>%
  mutate(pct = num_ads / sum(num_ads)) %>%
  ungroup()

```{r fig.height=7}
ggplot(df_brand_country, aes(x = pct, y = reorder_within(brand, pct, country_full), fill = brand)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ country_full, ncol = 2, scales = "free_y") +
  scale_x_continuous(labels = percent_format()) +
  scale_y_reordered() +
  labs(
    title = "Top 10 Brands - Market Share by Country",
    x = "Marketshare",
    y = NULL
  ) +
  theme_minimal()
```

### Top 5 Models for Each of the 10 Largest Brands

```{r}
df_model_brand <- df %>%
  filter(brand %in% top_brands) %>%
  group_by(brand, model) %>%
  summarise(num_ads = sum(num_ads), .groups = "drop") %>%
  arrange(brand, desc(num_ads)) %>%
  group_by(brand) %>%
  slice_head(n = 5) %>%
  ungroup()

```{r fig.height=7}
ggplot(df_model_brand, aes(x = num_ads, y = fct_reorder(model, num_ads), fill = brand)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ brand, ncol = 2, scales = "free_y") +
  scale_x_continuous(labels = comma) +
  labs(
    title = "Top 5 Models for Each of the 10 Largest Brands",
    x = "Number of Ads",
    y = NULL
  )
```